\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-10-27 metatheory-condensed.tex
explain substitution

\TT is intended as a foundational model. So unlike our programmer-facing plugin,
typing is structural rather than nominal, and
it supports family polymorphism by providing MLTT-style typing rules that
can be used to \emph{encode} core aspects of the language supported by our
plugin, rather than building them in directly.
%
For example, \TT supports the extension of inductive types through
formation rules that enable the encoding of the further binding of W-types~\cite{martin1984intuitionistic} in linkages.
Also unlike our plugin, \TT does not feature automations such as
the generation of propositional equalities and partial recursors,
but it is expressive enough to encode their definitions.

@@

\noindentparagraph{Syntax.}

>>>>

\TT is intended as a foundational model. So unlike our programmer-facing plugin,
typing is structural rather than nominal, and
it supports family polymorphism by providing MLTT-style typing rules that
can be used to \emph{encode} core aspects of the language supported by our
plugin, rather than building them in directly.
%
For example, \TT supports extending inductive types through
formation rules that enable the encoding of the further binding of W-types~\cite{martin1984intuitionistic} in linkages.
Also unlike our plugin, \TT does not feature automations such as
the generation of propositional equalities and partial recursors,
but it is expressive enough to encode their definitions.

@@

\noindentparagraph{MLTT.}

\cref{fig:typing-selected} presents the syntax of \TT.
The new forms (relative to MLTT) are highlighted.

As is standard in MLTT presentations, we use explicit substitutions~\cite{substcalculus,abadi1989subst}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
of the syntax rather than meta-operations.
%
Variables are represented by de Bruijn indices:
\var{n} is the variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{(\lam{\var{1}})}$.
%
Formation rules for substitutions have the form
$\goodSub{\Gm}{\gm}{\Dl}$, which says that applying~$\gm$ to terms
valid in the context~$\Dl$ yields terms valid in~$\Gm$.
%
The two main forms of substitutions are weakening and extension:
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/proj} states that if $t$ is a dependent pair
of type $\TySigma{A}{B}$, then $\snd{t}$ has type $\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$.

%Note that we are actually using predicative universe, so we actually have different levels of universes $\cU_i$. 
%We omit universe levels in our latex presentation for conciseness and readability.
